---
title: 'Geog 4/6300: Lab 3'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Confidence intervals and sampling

**Your name: {Name here}**

**Overview:**

In this lab, we will be calculating estimated food insecurity at state level. We will be using individual level “microdata” from the Current Population Survey (CPS). It is designed as an ongoing (collected monthly) set of data on financial and demographic characteristics. One main use of the CPS is to calculate national levels of food insecurity. Each December, a food security supplement is added to the regular survey, and data from the supplement is included here.

To load these data, load the csv file, which is included in the lab folder:

```{r message=FALSE}
cps_data<-read_csv("data/IPUMS_CPS_FoodSec.csv")
```

This contains a csv file with microdata from the CPS that is de-identified and publically available through the Minnesota Population Center (https://cps.ipums.org/cps/index.shtml). There is also a codebook that is part of the Github repo (IPUMS_CPS_CODEBOOK.pdf) describing each of those variables.

###Part 1: Calculating national food insecurity###
For this lab, you will be using the FSSTATUS variable, which describes the food security of respondents. While food security status is often grouped into “low” and “very low” food security, these two are often just combined to a single measure: food insecure. The codebook (see the link above) lists the values assocated with missing or "Not in universe" (NIU) records. For the purposes of this lab, an additional column called FSSTATUS_cat has been added with text versions of the FSSTATUS classification: food_secure, low_fs, and "verylow_fs".

**Question 1** _Filter out all records in this dataset that have missing or NIU observations for food insecurity. The remaining entries will be labeled as 1, 2, or 3. Create a new variable that provides appropriate text labels based on how they are described in the codebook. The `case_when` function may be particularly helpful._

```{r}
#Code goes here
```

**Question 2** _Then use group_by and summarise to calculate the number of individuals grouped in each of your new food security categories.  Call this table using kable. Based on these results, calculate the estimated national food insecurity rate (those not described as food secure). Make sure that result is visible either in your code chunk or in text below it._

```{r}
#Code goes here
```

**Question 3** _Using the formula for confidence intervals for proportions shown in class, calculate a confidence interval for the rate you identified in question 1. Make sure that confidence interval is shown in the knitted document you create. Interpret what that confidence interval tells you--how would you explain it to a friend not in this class?_

```{r}
#Code goes here. 
```

{Interpretation goes here.}

### Part 2: Analyzing state food insecurity data

We can also use the "STATE" variable to calculate rates for each state. You can adapt the code from question 1 above to include STATE as a second grouping varirable. 

**Question 4** _Create an estimated food insecurity rate for each state from these data. To do so, you'll need to (1) create counts for each response (food secure, low food security, very low food insecurity), and (2) transform the data so all three are spread out in wide format. You can then (3) sum the two variables for food insecurity and divide by the total responses within each state to get a percentage. Call the head of your table when done using kable._

```{r}
#Code goes here.
```

**Question 5** _Calculate the error term (the standard error * the z score for 95% confidence) for each state using the equation for standard error of proportions. Remember that each state will have a different sample size, which will effect this equation. Call the head of your table when done._

```{r}
#Code goes here
```

**Question 6** _Create a graph of the confidence intervals from question 4 using ggplot, geom_linerange (for the CI range), and geom_point (for the estimated mean) as shown in class. Arrange the states along the y axis based on the estimated food insecurity rate. See the code shared in class lectures for a template to work from._ 

```{r}
#Code goes here
```

**Question 7** _How would you interpret what this graph shows about food insecurity nationally? Write a paragraph that is appropriate for a general audience, describing the overall pattern of food insecurity rates and any regional or other trends that you notice. Use the table you created in question 4 to give specific figures that illustrate these points._

{Explanation goes here}

**Question 8** _Compare the margin of error (error term) you calculated for Georgia to the national margin of error. How do they differ? Mathematically, why are they different?_

{Explanation goes here.}

**Challenge question**
The lab data folder includes two additional files:
* A geotiff raster showing proximity (in unspecified units) to SNAP-authorized supermarkets in the Atlanta metro: atl_snapsupers_proxmity.tig
* A polygon layer of census tracts in the core Atlanta metro along with their population: atl_tracts.gpkg
* A point layer of park spaces listed on OpenStreetMap: atl_osm_parks

Using the techniques discussed in the scripts on spatial operations and raster analysis, map out the census tracts and/or parks located in areas with the lowest proximity (highest values on the raster) to supermarkets. Make sure your map(s) is visible in the output and clearly show the tracts/parks where new food options could be improve the proximity of healthy foods. Provide an interpretation of your map(s) that explains what patterns they show.

```{r}
#Code goes here
```

(Your interpretation here)

**Disclosure of assistance:** _Besides class materials, what other sources of assistance did you use while completing this lab? These can include input from classmates, relevant material identified through web searches (e.g., Stack Overflow), or assistance from ChatGPT or other AI tools. How did these sources support your own learning in completing this lab?_


**Lab reflection:** _How do you feel about the work you did on this lab? Was it easy, moderate, or hard? What were the biggest things you learned by completing it?_

